Data Preparation

### 2019-01-14 Data Preparation.Rmd
bug when extracting overview fixed

# Variable Engineering
##Program Data
The analysis performed on the course data is aimed at discoving what the contribution of each course is towards the fullfilment of the AoDs, and comparing different types of curricula, or programs, based on this infromation.

@@

```{r extracting overview, cache = TRUE}
# Setup
n_catalogue <- length(corpus_catalogues)
calendar_years <- c("2014-2015", "2015-2016", "2016-2017", "2017-2018", "2018-2019")
d_description <- tibble(Code            = character(0),
                        `Calendar Year` = character(0),
                        Overview        = character(0))

@@

    # Extract description
    overview <- if(page + 1 %in% first_pages_overview) cat_overview[page]
                else paste(cat_overview[page : (page + 1)], collapese = " ")

    # Save overview
    Code <- substr(overview, start = 1, stop = 7)
    year <- calendar_years[n]
    d_description <- d_description %>%
      add_row(
        Code = Code,
        `Calendar Year` = year,
        Overview = overview
      )

@@

```{r extracting description, cache = TRUE}
# Setup
symbol_to_line <- c("\r\n", "\n") %>%
  paste(collapse = "|")
start_descrip <- c("^Description", "^Course Description") %>%
  paste(collapse = "|")
end_descrip   <- c("^Literature$", "^Recommended Literature", "^ Required Litera",
                   "^Required Litera", "^ Literature$", "Literature \\(all") %>%
  paste(collapse = "|")
d_description <- d_description %>%
  mutate(Description = NA)
# Loop
for(course in 1 : nrow(d_description)){

@@

Separately, we want to add the manual textual information to`d_course`which contains information at the course-level such as in which cluster and concentration(s) they belong, and in which period(s) they are offered.
```{r join manuals to d_course}
d_course <- full_join(d_course, d_manual, by = "Code")
```

@@

To put everything into tidy text format, we first create three tibbles `d_overview_tidy`, `d_description_tidy`, `d_manual` that respectively store the overviews, descriptions and text from manuals in the tidy text format, with one row per course-year-word (and course-word for `d_manual`).
```{r tidy d_description}
d_overview <- d_description %>%
  select(Code, `Calendar Year`, Overview) %>%
  unnest_tokens(output = word, input = Overview)
d_description <- d_description %>%
  select(Code, `Calendar Year`, Description) %>%
  unnest_tokens(output = word, input = Description)
d_manual <- unnest_tokens(d_manual, output = word, input = Manual)
```

### Stemming
We then use the function `hunspell__stem`, which returns valid stems for a given word, to stem the words in the tibbles `d_overview_tidy` and `d_description_tidy`.

We first create a function `stem_hunspell` which, given a term, returns its most basic stem (the last one from the list of stem returned by `hunspell__stem`). We would like to apply `stem_hunspell` on the words of `d_overview_tidy` and `d_manual`, since they have similar structures, we row bind them into `dictionary` to ease the application of `stem_hunspell`. However, since, `stem_hunspell` is not a vectorized function and the number of words in `dictionary` is large, we first use distinct on the `word` variable to find a list containing *once* each term present in the course overviews and manuals. Then, we use `mutate` to apply the function `stem_hunspell` on each word from our dictionary and save its results as a new column `word_stem`.
```{r stemming dictionary, cache = TRUE}
stem_hunspell <- function(term) {
  stems   <- hunspell_stem(term)[[1]] # look up the term in the dictionary
  n_stems <- length(stems)            # number of stems
  if (n_stems == 0) term              # if no stems, return original term
  else              stems[[n_stems]]  # if multiple stems, return last (most basic) stem
}
dictionary <- select(d_overview, Code, word) %>%
  rbind(d_manual) %>% 
  select(word) %>%
  distinct %>%
  mutate(word_stem = purrr::map_chr(.x = word,
                                    .f = stem_hunspell))
# terms for which the stem differs from the original word
filter(dictionary, word != word_stem)
```
Finally, we create a function `stem_with_dictionary` that performs a left join on the dataframe it takes as input with `dictionary`, thus adding the word stems to the original dataframe.
Then, we use `stem_with_dictionary` to include the stems of all words in `d_description`, `d_overview`, and `d_manual`.
```{r stemming data sets}
stem_with_dictionary <- function(data) data %>%
                                         left_join(dictionary, by = "word") %>%
                                         rename(word_original = word,
                                                word = word_stem)

d_description <- stem_with_dictionary(d_description)
d_overview <- stem_with_dictionary(d_overview)
d_manual <- stem_with_dictionary(d_manual)
rm(stems, stem_hunspell, stem_with_dictionary)
print(d_overview) # See humanities - humanity
```

>>>>

# Variable Engineering
## Program Data
The analysis performed on the course data is aimed at discoving what the contribution of each course is towards the fullfilment of the AoDs, and comparing different types of curricula, or programs, based on this infromation.

@@

```{r extracting overview, cache = TRUE}
# Setup
n_catalogue <- length(corpus_catalogues)
calendar_years <- c("2014-2015", "2015-2016", "2016-2017", "2017-2018", "2018-2019")
d_description <- tibble(Code            = character(0),
                        `Calendar Year` = character(0),
                        Overview        = character(0),
                        Description     = NA)

@@

    # Extract description
    overview <- if((page + 1) %in% first_pages_overview) cat_overview[page]
                else paste(cat_overview[page : (page + 1)], collapse = " ")

    # Save overview
    Code <- substr(overview, start = 1, stop = 7)
    year <- calendar_years[n]
    d_description <- d_description %>%
      add_row(
        Code = Code,
        `Calendar Year` = year,
        Overview = overview,
        Description = NA
      )

@@

```{r extracting description, cache = TRUE}

# Setup
symbol_to_line <- c("\r\n", "\n") %>%
  paste(collapse = "|")
start_descrip <- c("^Description", "^Course Description") %>%
  paste(collapse = "|")
end_descrip   <- c("^Literature$", "^Recommended Literature", "^ Required Litera",
                   "^Required Litera", "^ Literature$", "Literature \\(all") %>%
  paste(collapse = "|")

# Loop
for(course in 1 : nrow(d_description)){

@@

@@

To put everything into tidy text format, we first create three tibbles `d_overview_tidy`, `d_description_tidy`, `d_manual` that respectively store the overviews, descriptions and text from manuals in the tidy text format, with one row per course-year-word (and course-word for `d_manual`).
```{r tidy d_description}
d_overview <- d_description %>%
  select(Code, `Calendar Year`, Overview) %>%
  unnest_tokens(output = word, input = Overview)

d_description <- d_description %>%
  select(Code, `Calendar Year`, Description) %>%
  unnest_tokens(output = word, input = Description)

d_manual <- unnest_tokens(d_manual, output = word, input = Manual)
```

### Stemming
We then use the function `hunspell_stem`, which returns valid stems for a given word, to stem the words in the tibbles `d_overview_tidy` and `d_description_tidy`.

We first create a function `stem_hunspell` which, given a term, returns its most basic stem (the last one from the list of stem returned by `hunspell_stem`). We would like to apply `stem_hunspell` on the words of `d_overview_tidy` and `d_manual`, since they have similar structures, we row bind them into `dictionary` to ease the application of `stem_hunspell`. However, since, `stem_hunspell` is not a vectorized function and the number of words in `dictionary` is large, we first use distinct on the `word` variable to find a list containing *once* each term present in the course overviews and manuals. Then, we use `mutate` to apply the function `stem_hunspell` on each word from our dictionary and save its results as a new column `word_stem`.
```{r stemming dictionary, cache = TRUE}
stem_hunspell <- function(term) {
  stems   <- hunspell_stem(term)[[1]] # look up the term in the dictionary
  n_stems <- length(stems)            # number of stems
  if (n_stems == 0) term              # if no stems, return original term
  else              stems[[n_stems]]  # if multiple stems, return last (most basic) stem
}

dictionary <- select(d_overview, Code, word) %>%
  rbind(d_manual) %>% 
  select(word) %>%
  distinct %>%
  mutate(word_stem = purrr::map_chr(.x = word,
                                    .f = stem_hunspell))
# terms for which the stem differs from the original word
filter(dictionary, word != word_stem)
```
Finally, we create a function `stem_with_dictionary` that performs a left join on the dataframe it takes as input with `dictionary`, thus adding the word stems to the original dataframe.

Then, we use `stem_with_dictionary` to include the stems of all words in `d_description`, `d_overview`, and `d_manual`.
```{r stemming data sets}
stem_with_dictionary <- function(data) data %>%
                                         left_join(dictionary, by = "word") %>%
                                         rename(word_original = word,
                                                word = word_stem)

d_description <- stem_with_dictionary(d_description)
d_overview <- stem_with_dictionary(d_overview)
d_manual <- stem_with_dictionary(d_manual)
rm(stem_hunspell, stem_with_dictionary)
print(d_overview) # See humanities - humanity
```
